\begin{acknowledgment}
	The researchers would like to express their deepest gratitude to God Almighty for granting them strength, wisdom, and perseverance to complete this study.
	
	The researchers extend their heartfelt appreciation to their thesis adviser, Dr. Valentine Blez Lampayan, for her unwavering support, guidance, and valuable insights throughout the development of this thesis. Her expertise and encouragement have played an important role in refining the study. A special thanks to Professor Marrick Neri for sharing his insights and expertise during the researchers’ visit to the University of the Philippines Diliman. His knowledge and recommendations have significantly contributed to the improvement of the thesis. The researchers are also deeply grateful to Leah Manongas, whose work served as a reference for the models used in this study. Her thesis provided a strong foundation that helped the researchers in the shelter location allocation process. 
	
	% mandatory acknowledgement to the creator of the bulsu thesis template files
	% do no delete
	Special thanks to the creator of bulsuthsis.cls, and its accompanying files thesis.tex and thesis-structure.tex, Mr. Harris R. Dela Cruz. With these, the authors were able to typeset conveniently and efficiently an elegant and beautiful thesis.
	% do no delete
	% mandatory acknowledgement to the creator of the bulsu thesis template files
	
	The researchers are also deeply grateful to the IT Experts, MDRRMO, and MSWDO Staff of Calumpit who took the time to evaluate our system and provide valuable feedback. Their insights have helped us identify areas for improvement and refine our work. 
	
	The researchers would like to express sincere gratitude to the Department of Science and Technology (DOST) for their financial support, which greatly contributed to the completion of this thesis. Their assistance provided the group with resources and opportunities necessary to pursue and accomplish this research.
	
	Furthermore, the researchers extend their appreciation to their friends and one of the researcher’s significant other for their continuous support, understanding, and patience as well as their encouragement which have been helpful in keeping their mental well-being and became our source of motivation throughout this journey.
	
	Lastly, the researchers would like to express their utmost gratitude to their families, whose love and support have been their greatest strength. Their belief in the researchers and their encouragement throughout this academic journey have been invaluable.
	
	This research would not have been possible without the support and contributions of everyone mentioned. Thank you all for being part of this journey.
	
	\begin{flushright}
	Bryan Jett T. Calulo\\
	Lovely Angeline OL. Cunanan\\
	Elijah Iñigo C. Fabian\\
	\end{flushright}

\end{acknowledgment}
